Database comparison

### 2025-03-28 2.Database_Comparison.Rmd
Updated database comparison

---
title: "Code used to generate paper results : Database comparison and CT characteristics"
author: "Julie Devis"
date: "2023-08-28"
output: 
  html_document:
    theme: cosmo
    df_print : paged
    highlight : default
    toc: yes
    number_sections : TRUE
    toc_depth : 4
---

@@

```{r load_data, warning=FALSE, message=FALSE}
load("../CTdata/eh_data/all_genes.rda")
load("../CTdata/eh_data/CT_genes.rda")

CT_and_CTP_genes <- CT_genes
CT_genes <- (filter(CT_genes, CT_gene_type == "CT_gene"))

@@

CT lists from other databases have been checked (using GTEx and our 
`GTEx_expression()` funtion and GeneCards) in order to remove duplicated
gene names or deprecated ones and allow comparison between databases.

@@

lost[lost$external_gene_name == "C21orf59", "Gene_Name"] <- "CFAP298"
lost[lost$external_gene_name == "C11orf57", "Gene_Name"] <- "NKAPD1"
lost[lost$external_gene_name == "C7orf55", "Gene_Name"] <- "FMC1"
lost[lost$external_gene_name == "C10orf12", "Gene_Name"] <- "LCOR"
lost[lost$external_gene_name == "RPL19P12", "Gene_Name"] <- "RPL19P12"
lost[lost$external_gene_name == "C16orf59", "Gene_Name"] <- "TEDC2"
lost[lost$external_gene_name == "TTTY15", "Gene_Name"] <- "USP9Y"
lost[lost$external_gene_name == "C17orf53", "Gene_Name"] <- "HROB"
lost[lost$external_gene_name == "C1orf112", "Gene_Name"] <- "FIRRM"
lost[lost$external_gene_name == "C12orf66", "Gene_Name"] <- "KICS2"
lost[lost$external_gene_name == "C9orf84", "Gene_Name"] <- "SHOC1"
lost[lost$external_gene_name == "C10orf25", "Gene_Name"] <- "ZNF22-AS1"
lost[lost$external_gene_name == "C20orf197", "Gene_Name"] <- "LINC02910"
lost[lost$external_gene_name == "C3orf67", "Gene_Name"] <- "CFAP20DC"
lost[lost$external_gene_name == "C8orf37", "Gene_Name"] <- "CFAP418"
lost[lost$external_gene_name == "C22orf34", "Gene_Name"] <- "MIR3667HG"
lost[lost$external_gene_name == "C9orf131", "Gene_Name"] <- "SPATA31G1"

@@

external_names_to_keep <- gene_synonym %>%
   filter(external_synonym %in% missing_Bruggeman) %>%
   filter(external_gene_name %in% all_genes$external_gene_name) %>% 
   mutate(Gene_Name = external_gene_name)

@@

length(Bruggeman_CT$Gene_Name[!Bruggeman_CT$Gene_Name %in% 
                                all_genes$external_gene_name])
```

38 genes are not found with their names in all genes

@@

transcript_prob <- lost %>% 
  filter(testis_specificity == "testis_specific" |
           testis_specificity == "testis_preferential") %>%
  filter(TCGA_category == "activated" | TCGA_category == "multimapping_issue") %>% 
  filter(CCLE_category == "activated") %>% 
  dim()
```

>>>>

---
title: "Database comparison"
output: 
  html_document:
    theme: cosmo
    df_print : paged
    highlight : default
    toc: yes
    number_sections : TRUE
    toc_depth : 4
---

@@

```{r load_data, warning=FALSE, message=FALSE}
CT_and_CTP_genes <- CT_genes()
CT_genes <- (filter(CT_genes(), CT_gene_type == "CT_gene"))

@@

CT lists from other databases have been checked (using GTEx and our 
`GTEx_expression()` function and GeneCards) in order to remove duplicated
gene names or deprecated ones and allow comparison between databases.

@@

lost[lost$external_gene_name == "C21orf59", "Gene_Name"] <- "CFAP298"
lost[lost$external_gene_name == "C11orf57", "Gene_Name"] <- "NKAPD1"
lost[lost$external_gene_name == "C7orf55", "Gene_Name"] <- "FMC1"
lost[lost$external_gene_name == "C10orf12", "Gene_Name"] <- "LCOR"
lost[lost$external_gene_name == "RPL19P12", "Gene_Name"] <- "RPL19P12"
lost[lost$external_gene_name == "C16orf59", "Gene_Name"] <- "TEDC2"
lost[lost$external_gene_name == "TTTY15", "Gene_Name"] <- "USP9Y"
lost[lost$external_gene_name == "C17orf53", "Gene_Name"] <- "HROB"
lost[lost$external_gene_name == "C1orf112", "Gene_Name"] <- "FIRRM"
lost[lost$external_gene_name == "C12orf66", "Gene_Name"] <- "KICS2"
lost[lost$external_gene_name == "C9orf84", "Gene_Name"] <- "SHOC1"
lost[lost$external_gene_name == "C10orf25", "Gene_Name"] <- "ZNF22-AS1"
lost[lost$external_gene_name == "C20orf197", "Gene_Name"] <- "LINC02910"
lost[lost$external_gene_name == "C3orf67", "Gene_Name"] <- "CFAP20DC"
lost[lost$external_gene_name == "C8orf37", "Gene_Name"] <- "CFAP418"
lost[lost$external_gene_name == "C22orf34", "Gene_Name"] <- "MIR3667HG"
lost[lost$external_gene_name == "C9orf131", "Gene_Name"] <- "SPATA31G1"
lost[lost$external_gene_name == "C17orf80", "Gene_Name"] <- "MTNAP1"

@@

external_names_to_keep <- gene_synonym %>%
   filter(external_synonym %in% missing_Bruggeman) %>% # take those in Bruggeman that are synonym
   filter(external_gene_name %in% all_genes$external_gene_name) %>% # but that are actually in all_genes with another name
   mutate(Gene_Name = external_gene_name)

@@

length(Bruggeman_CT$Gene_Name[!Bruggeman_CT$Gene_Name %in% 
                                all_genes$external_gene_name])

39 genes are not found with their names in all genes

@@

transcript_prob <- lost %>% 
  pull(IGV_backbone) %>% 
  table()
```
